### 2015-05-02 draft.tex
..

\begin{figure}
\begin{lstlisting}
f(pkt):
  if RO.in_range(pkt, src, 0, 10) then
     if RO.in_range(pkt, dst, 0, 10) then
        RO.deliver()
     else 
        RO.add(pkt, (dst, src), trusted);
        RO.deliver()
  else if (RO.in_set(pkt, (src,dst), trusted) then
         RO.deliver()
       else
         RO.drop
       end
  end                
  \end{lstlisting}
\label{fig:firewallcode}
\caption{A Stateful Firewall}
\end{figure}

@@

   \begin{figure}
\begin{lstlisting}
Module RO

Type pkt (* contains a range for each field *)

Type field (* enum each field *)

\caption{Signature of the packet portion of the API}
\end{figure}

@@

@@

\begin{lstlisting}
  compile (f: packet -> forwarding_decision) : policy =
      push most general packet to (empty) stack
      set the current location to be the root of an empty decision tree
      while (stack is not empty)
         pop the top of the stack as pkt
         run f of pkt resulting in a forwarding decision fd (and a modified packet pkt)
         record the forwarding decision for the symbolic packet pkt at the current location in the decision tree
         set the current location to be the root of the decision tree and repeat

@@

>>>>

\begin{figure}
\begin{lstlisting}
f(pkt):
  if Oak.in_range(pkt, src, 0, 10) then
     if Oak.in_range pkt dst  [(0,10)] then
        Oak.deliver
     else 
        Oak.add pkt (Oak.IpDst, Oak.IpSrc) ``conn'';
        Oak.deliver
  else if Oak.in_rel pkt (Oak.IpSrc,Oak.IpDst) ``conn'' then
         Oak.deliver
       else
         RO.drop
       end
  end                
  \end{lstlisting}
\label{fig:firewallcode}
\caption{A Stateful Firewall}
\end{figure}

@@

@@

\begin{figure}
  \begin{lstlisting}
(* a standard packet field *)
type field = IpSrc | IpDst

(* range is a union of intervals *)
type range = (int*int) list

(* forwarding decision can deliver, drop, forward, multicast *)
type forwarding_decision =
	| Deliver
	| Drop
	| ForwardTo of int
	| Multicast of forwarding_decision * forwarding_decision
	| Ctrl

(* abstract packet type *)
type packet

(* minimum, maximum value for a range interval *)
val min_value: int 
val max_value: int

(* in_range p f r returns true if the value of field f of packet p is in range r, and false otherwise *)
val in_range: packet -> field -> range -> bool
\end{lstlisting}

\caption{Signature of the packet and range-branching portion of the API}
\end{figure}

\begin{figure}
  \begin{lstlisting}
(* name of a relation *)
type relation = string

(* add packet's fields to a relation *)
val add: packet -> field list -> relation -> unit

(* remove packet's fields from a relation *)
val remove: packet -> field list -> relation -> unit

(* test if a packet's fields are in a relation *)
val in_relation: packet -> field list -> relation -> bool
\end{lstlisting}

\begin{figure}
  \begin{lstlisting}
(* compile a user-function into a policy *)
val compile: (packet -> forwarding_decision) -> policy

(* run the user-defined policy with a set of input packets *)
val run: policy -> ((field * int) list) list  -> unit

(* Print the policy to stdout *)
val print_policy: policy -> unit
  \end{lstlisting}
  \caption{Signature of the interactive portion of the API}
\end{figure}

@@

\begin{figure}
\begin{lstlisting}
let compile (f: packet -> forwarding_decision) : policy =
      push most general packet to (empty) stack
      set the current location to be the root of an empty decision tree
      while (stack is not empty)
         pop the top of the stack as pkt
         run f of pkt resulting in a forwarding decision fd (and a modified packet pkt)
         record the forwarding decision for the symbolic packet pkt at the current location in the decision tree
         set the current location to be the root of the decision tree and repeat

\end{lstlisting}
\caption{Pseudocode for the compile function}
\end{figure}

@@

\begin{figure}
\begin{lstlisting}
let run (pol: policy) (inputs: ((field*int) list) list) : unit =
      build the symbolic rules for policy pol
      generate the concrete rules with all relations empty
      for each packets in input,
          find the first of the current rule matching the packet, and record the forwarding decision fd
          if packet is sent to the controller then
              evaluate the policy on the packet, potentially adding or removing tuples from relations, record forwarding decision fd'
              regenerate concrete rules
              forward the packet according the forwarding decision fd'
          else
              forward the packet according to fd 
\end{lstlisting}

\caption{Pseudocode for the run function}
  \end{figure}
